Typesafe Coordinate Systems in High-Throughput Sequencing Applications

### 2022-09-13 manuscript/main.tex
Remove \filbreak (2) and revise footnote

Instead, consider strongly and statically typed languages with generic and polymorphic features. Define a type class \texttt{Interval} of which there are four types, corresponding to ZBHO, ZBC, OBHO, and OBC coordinate systems, and for which there exists a function, \textit{to}, (or conversely, \textit{from}) that maps between constituent types. Note that the types may be freely interconverted, as there exists a bijective mapping between any two coordinate systems (Figure \ref{fig:transitiongraph}). While these types could all be defined individually, one might instead leverage polymorphic features of the type system, especially in systems where a type can be parameterized on a constant value. 
\filbreak
In C++, D, or similar languages this could be done by templating a single \texttt{Interval} class or struct on some enums defining basis and openness:
\begin{verbatim}
    // C++
    Interval <CoordSystem::zbho> i (0, 100);
    // D
    auto i = Interval!CoordSystem.zbho(0, 100);
    // Rust
    let i = Interval::<ZeroBased, HalfOpen>::from_int(0, 100);
\end{verbatim}

@@

To instead create a function \texttt{fun} that accepts input of all coordinate systems, the function must be polymorphic or generic over the coordinate(s) type. Then in use, \texttt{fun} must be instantiated\footnote{In C++, D, Rust the function would undergo monomorphization; in e.g. Haskell or OCaml not} with the desired coordinate system type and thereby be parameterized by a known \texttt{Coordinates} type, so type-safety is still achieved. In D this can be inferred, so the function will be automatically instantiated with the correct type. The coordinate system provides compile-time (via the type system) and run-time checks (bounds) to ensure coordinates are valid for their coordinate system; \textit{i.e.}, a one-based system must not contain 0 as a coordinate while a half-open coordinate system’s end coordinate must be greater than its start coordinate.

@@

\filbreak
A pseudocode excerpt is shown below.\\

>>>>

Instead, consider strongly and statically typed languages with generic and polymorphic features. Define a type class \texttt{Interval} of which there are four types, corresponding to ZBHO, ZBC, OBHO, and OBC coordinate systems, and for which there exists a function, \textit{to}, (or conversely, \textit{from}) that maps between constituent types. Note that the types may be freely interconverted, as there exists a bijective mapping between any two coordinate systems (Figure \ref{fig:transitiongraph}). While these types could all be defined individually, one might instead leverage polymorphic features of the type system, especially in systems where a type can be parameterized on a constant value.

In C++, D, or similar languages this could be done by templating a single \texttt{Interval} class or struct on some enums defining basis and openness:
\begin{verbatim}
    // C++
    Interval <CoordSystem::zbho> i (0, 100);
    // D
    auto i = Interval!CoordSystem.zbho(0, 100);
    // Rust
    let i = Interval::<ZeroBased, HalfOpen>::from_int(0, 100);
\end{verbatim}

@@

To instead create a function \texttt{fun} that accepts input of all coordinate systems, the function must be polymorphic or generic over the coordinate(s) type. Then in use, \texttt{fun} must be instantiated\footnote{Ideally via monomorphization as in C++, D, or Rust for performance.} with the desired coordinate system type and thereby be parameterized by a known \texttt{Coordinates} type, so type-safety is still achieved. In D this can be inferred, so the function will be automatically instantiated with the correct type. The coordinate system provides compile-time (via the type system) and run-time checks (bounds) to ensure coordinates are valid for their coordinate system; \textit{i.e.}, a one-based system must not contain 0 as a coordinate while a half-open coordinate system’s end coordinate must be greater than its start coordinate.

@@

A pseudocode excerpt is shown below.\\
